### 2020-04-08 manuscript/manuscript.RmD
- ms: updates on methods, results, and discussion section

---
title             : "How Do Like and Dislike Buttons Affect Communication? A Privacy Calculus Approach to Understanding Self-Disclosure Online in a One-Week Field Experiment"
shorttitle        : "Popularity cues and the privacy calculus"
author: 
  - name          : "Author"
    affiliation   : "1"
    corresponding : yes
    address       : "Postal address"
    email         : "my@email.com"
affiliation:
  - id            : "1"
    institution   : "Wilhelm-Wundt-University"
authornote: |
abstract: |
    According to the privacy calculus, both privacy concerns and expected gratifications explain self-disclosure online. So far, most findings were based on self-reports, and little is known about whether the privacy calculus can explain also observations of authentic behavior. Likewise, we still know comparatively little as to whether the privacy calculus can be influenced by the design of online websites, including for example popularity cues such as like and dislike buttons. To answer these questions, we ran a preregistered one-week field experiment. Participants were randomly distributed to three different websites, on which they could discuss a current political topic. The final sample consisted of `r knitr::load_cache(label = "load-data", object = "n_matched")` participants. Although the originally preregistered model could not be confirmed, the results showed that the privacy calculus variables predicted a considerable share of actual self-disclosure. The impact of popularity cues was negligible. In conclusion, the results indicate that self-disclosure online can be explained by privacy concerns and specific gratifications. This finding has several implications. For example, it provides further evidence against the privacy paradox.
keywords          : "privacy calculus, self-disclosure, popularity cues, field experiment, structural equation modeling, preregistration"
bibliography      : 
  - [../bibliography/r-references.bib]
  - [../bibliography/references.bib]
floatsintext      : yes
figsintext        : yes
figurelist        : no
tablelist         : no
footnotelist      : no
linenumbers       : yes
mask              : yes
wordcount         : "`r wordcountaddin::word_count('../manuscript/manuscript.Rmd')`"
draft             : no
documentclass     : "apa6"
classoption       : "man"
output            : papaja::apa6_pdf
# output            : papaja::apa6_word
header-includes:
  - \setlength{\parskip}{0em}
  - \raggedbottom
---

@@

Understanding why people disclose personal information online remains a critical question for both society and academic research. Originally, self-disclosure online was thought to be mostly erratic---for example, it was assumed that self-disclosure cannot be predicted by assessing people's personal beliefs, concerns, or standpoints. Most prominently, the privacy paradox stated that people self-disclose vast amounts of personal information online *despite* having substantial concerns about their privacy [@barnesPrivacyParadoxSocial2006; @taddickenUsesPrivacyOnline2011].

Somewhat surprisingly, despite its popularity in the media [@newyorkpublicradioPrivacyParadox2018] the privacy paradox has garnered little empirical support. A recent meta-analysis revealed that the correlation between privacy concerns and self-disclosure on SNS is *r* = -.13 [@baruhOnlinePrivacyConcerns2017], which indicates that privacy concerns are indeed related to self-disclosure online.

Rather than further pursuing the privacy paradox, a large share of current day research hence posits that self-disclosure online can be explained---at least partly---by means of the so-called _privacy-calculus_ [@krasnovaOnlineSocialNetworks2010]. 
Building on the work of @lauferPrivacyConceptSocial1977, the privacy calculus  claims that both expected risks *and* expected benefits explain self-disclosure. Specifically, by operationalizing expected risks as privacy concerns, several studies have shown that experiencing greater privacy concerns is related to disclosing less information [@heirmanPredictingAdolescentsDisclosure2013; @koohikamaliInvestigationDynamicModel2019].

@@

Second, current research on the privacy calculus is often criticized for not explicitly focusing on the deliberation process of self-disclosure. 
According to critics [e.g., @knijnenburgDeathPrivacyCalculus2017], showing that concerns and gratifications both correlate with self-disclosure is not evidence for any substantial or explicit weighing of pros and cons. 
We agree and in this study hence explicitly focus on the privacy deliberation process itself. 
Moreover, and on a more general level, we aim to gauge the usefulness of further extending the privacy calculus model by adding new variables such as privacy deliberation, trust, and self-disclosure self-efficacy.

Finally, because the privacy calculus does not take place in a vacuum, and because it is often argued that self-disclosure can be easilty triggered by external circumstances, we analyze whether the privacy calculus can be affected by the design of a website. 
Specifically, in this study we investigate whether *popularity cues* such as like and dislike buttons have the power to affect the privacy calculus and to foster self-disclosure.

To test our research questions, we conducted a preregistered online field experiment, drawing from a representative sample of the German population. Participants were randomly distributed to one of three different websites, which either included only a like button, both a like and a dislike button, or no buttons at all. Over the course of one week participants had the chance to discuss a topical issue (i.e., prevention of terrorist attacks in Germany). Afterward, they answered our follow-up questionnaire with items pertaining to the privacy calculus variables.

@@

Being a primary means of regulating privacy [e.g., @masurSituationalPrivacySelfdisclosure2018], self-disclosure is our key variable of interest. 
There are two different understandings of self-disclosure in the literature: The first defines self-disclosure as _deliberate_ acts of sharing truthful information about the self with others [@jourardTransparentSelf1964]. 
The second considers _all_ acts of sharing information---whether active or passive, deliberate or unwitting---as self-disclosure, because each piece of information shared allows meaningful inferences to be made about a person [e.g., @watzlawickPragmaticsHumanCommunication2011]. 
In this paper we follow the latter approach, not least because recent years have vividly illustrated how it is possible to derive a plethora of insights about a person simply by analyzing his or her written communication [e.g., @kosinskiPrivateTraitsAttributes2013]. 
Moreover, independent from which position one chooses to adopt, it is possible to differentiate the content of self-disclosure into three different dimensions: breadth (i.e., number of topics covered), depth (i.e., intimacy of topics covered), and length (i.e., quantity of disclosure) [e.g., @omarzuDisclosureDecisionModel2000]. 
In this study we mainly focus on communication quantity, as we consider communication quantity to be a necessary precondition and hence valid proxy for self-disclosure.

Privacy concerns have been defined as follows: 
"Concerns about online privacy represent how much an individual is motivated to focus on his or her control over a voluntary withdrawal from other people or societal institutions on the Internet, accompanied by an uneasy feeling that his or her privacy might be threatened" [AUTHOR]. 
Previous research has found that people who are more concerned about their privacy than others are less inclined to share personal information [@baruhOnlinePrivacyConcerns2017; @dienlinPrivacyParadoxRelic2015; @heirmanPredictingAdolescentsDisclosure2013; @koohikamaliInvestigationDynamicModel2019].

@@

Although privacy concerns are related to self-disclosure, one can make the case that since most studies in the literature report only small effects, there should also be additional meaningful factors that contribute to explaining self-disclosure. Most prominently, it has been argued that people trade a loss of privacy for a gain in gratifications such as social capital, entertainment, information, or self-presentation [@ellisonNegotiatingPrivacyConcerns2011; @taddickenUsesPrivacyOnline2011]. By now, a large body of research has found support for this hypothesis [e.g., @krasnovaOnlineSocialNetworks2010; @minHowArePeople2015; @trepteCrossculturalPerspectivePrivacy2017].

@@

In the current literature on the privacy calculus there still seems to be a shortage of studies that explicitly analyze the decision process of actively comparing the pros and cons of disclosing information, although this point of criticism has been leveled several times [e.g., @knijnenburgDeathPrivacyCalculus2017] and although other fields such as behavioral economics have long focused on the underlying problem [e.g., @zhuPrivacyCalculusIts2017]. 
This criticism is justified.
The observation that both experiencing privacy concerns and expecting gratifications are related to self-disclosure does not bit itself necessitate an explicit weighing process 
Hence, we argue that the research on the privacy calculus would benefit significantly from analyzing this decision process explicitly. 
Building on @omarzuDisclosureDecisionModel2000 and @altmanPrivacyConceptualAnalysis1976, we hence address a novel concept that might best be termed _privacy deliberation_, which we define as the extent to which individual people explicitly compare positive and negative potential outcomes before communicating with others.

On the one hand, it seems plausible that deliberating about one's privacy would dampen subsequent self-disclosure, because refraining from regular communication---the primary means of connecting with others---requires at least a minimum of active and hence deliberate restraint. 
On the other hand, deliberating about one's privacy might also increase self-disclosure, as after having actively deliberated about the potential consequences, a person concerned about his or her privacy might arrive at the conclusion that in this situation self-disclosure is not only appropriate but expedient. 
In light of the paucity of empirical studies and the plausibility of both effects, we formulate the following research question:

@@

Several attempts have already been made to expand the privacy calculus [e.g., @dinevExtendedPrivacyCalculus2006]. 
Additional variables such as self-efficacy or trust have been introduced. 
Self-efficacy in the context of the privacy calculus captures whether people believe in their own capability to implement particular privacy behaviors in the future [@dienlinExtendedPrivacyCalculus2016].
These privacy behaviors can either refer to self-withdrawal (e.g., deleting inappropriate content) or self-disclosure (e.g., publishing a blog post).
Thus far, several studies have found that people who report more privacy self-efficacy also self-withdraw more online than others [e.g., @chenRevisitingPrivacyParadox2018].
In light of our focus on self-disclosure, in this study we investigate the influence of self-disclosure self-efficacy.

Trust can be conceptualized in two different ways [@gefenTrustTAMOnline2003]: It either captures "_specific_ beliefs dealing primarily with the integrity, benevolence, and ability of another party" [@gefenTrustTAMOnline2003, p. 55, emphasis added] or a "_general_ belief that another party can be trusted" [@gefenTrustTAMOnline2003, p.55, emphasis added]. 
Whereas specific trust beliefs focus on the causes of trust, general trust beliefs focus on the experience of trust.
@gefenTrustTAMOnline2003 prioritize specific trust beliefs (p. 60). 
In the online context, it is important to differentiate among several targets of trust [@sollnerWhyDifferentTrust2016]. 
Potential targets include (a) the information system, (b) the provider, (c) the Internet, and (d) the community of other users [@sollnerWhyDifferentTrust2016]. 
Trust plays a key role in online communication [@metzgerPrivacyTrustDisclosure2004]. 
For example, it has been demonstrated that people who put more trust in the providers of networks also disclose more personal information [@liEmpiricalStudiesOnline2011].

In conclusion, while we expect to find these relations as well, we would also like to determine whether the inclusion of all the other variables mentioned above, including the not yet researched concept of privacy deliberation, might potentially attenuate or even obviate the predictive capacity of self-efficacy and trust.

@@

@@

What is the effect of the communication context on the privacy calculus and on self-disclosure? 
First, it has often been noted that researchers should not exclusively focus on specific features of particular websites, for features are prone to change and quickly become obsolete [@foxDistinguishingTechnologiesSocial2017]. 
Instead, it has been suggested that researchers prioritize underlying latent structures, for example by analyzing what are know as affordances [e.g., @ellisonSocialNetworkSite2015; @foxDistinguishingTechnologiesSocial2017]. 
The concept of affordances was developed by @gibsonEcologicalApproachVisual2015, who argued that it is not the objective features of objects that determine behavior but rather subjective perceptions. 
Affordances are a mental representation of how a given entity might be used; as such, they are by definition subjective. 
There is much debate in the literature concerning what exactly defines an affordance [@evansExplicatingAffordancesConceptual2017]. 
For example, whereas @evansExplicatingAffordancesConceptual2017 propose three affordances for mediated communication (i.e., anonymity, persistence, and visibility), @foxDistinguishingTechnologiesSocial2017 suggest 10 affordances for SNSs alone (i.e., accessibility, bandwidth, social presence, privacy, network association, personalization, persistence, editability, conversation control, and anonymity).

As the privacy calculus states that both benefits and costs determine behavior, we suggest that popularity cues such as like and dislike buttons, which are categorized as "paralinguistic digital affordances" [@carrPredictingThresholdPerceived2018, p. 142], perfectly epitomize benefits and costs. 
The like button is positive; it expresses an endorsement, a compliment, a reward [@carrPredictingThresholdPerceived2018; @sumnerFunctionalApproachFacebook2017]. 
However, communication online is also often characterized by negative and critical debates [@ziegeleDynamicsOnlineNews2017]. 
As the dislike button is a major means of downgrading content it represents the cost and risk factor of the privacy calculus well. 
In fact, its stark negative effect might also explain why to date only a handful of major websites have implemented it (e.g., youtube, reddit or stackexchange).

Paralinguistic digital affordances and/or popularity cues have been shown to impact behavior [@kramerMasteringChallengeBalancing2020; @treptePrivacyCalculusContextualized2020].
For example, a large-scale field experiment in which 101,281 comments were analyzed found that comments with dislikes were more likely to receive further dislikes [@muchnikSocialInfluenceBias2013]. 
@stroudRecommendRespectAltering2017 demonstrated that when users had a different opinion than the one that was communicated in a post, they were more likely to click on a button labelled _respect_ compared to a button labelled _like_.

In this vein it seems plausible that popularity cues might also impact the privacy calculus [kramerMasteringChallengeBalancing2020].
First, on a primordial level, popularity cues serve as a means of reward and punishment, affecting behavior via instrumental conditioning [@skinnerScienceHumanBehavior2014]. 
Specifically, being complimented with a like should encourage future self-disclosure, while being punished with a dislike should inhibit future disclosure. 
Similarly, like buttons should be associated with being able to garner positive feedback, so implementing a like-button---similar to a compliment in the offline world---might leverage gratifications. Implementing a like or a dislike button might also bring people to more actively deliberate about whether or not it is actually worthwhile to disclose information. 
If both like and dislike buttons are present, privacy deliberation should increase even further. 
Finally, because people who are more concerned about their privacy are also more shy and risk averse [@dienlinPsychologyPrivacyAnalyzing2017], implementation of the dislike button should both stir privacy concerns and stifle self-disclosure. 
For a simplified overview of our theoretical model, see Figure \@ref(fig:model).

@@

@@

The online supplementary material (OSM) of this study include the data, research material, analyses scripts, and a reproducible version of this manuscript (see https://osf.io/hcqat/?view_only=5db35868738d40609b11e58cc343a9b0)
We preregistered the study using the registration form _OSF Prereg_, which includes hypotheses, sample size, materials, analyses, and exclusion criteria (see https://osf.io/a6tzc/?view_only=5d0ef9fe5e1745878cd1b19273cdf859). 
We needed to change our pre-defined plan in some cases. 
For a full account of all changes, see OSM. 
New analyses that were not preregistered appear in the section on exploratory analyses. 
For example, we also measured two additional variables that were not included in the preregistration (e.g., _specific_ gratifications and _general_ trust; see below), which are included in the exploratory analyses.

@@

The study was designed as an online field experiment with three different groups. 
The first group interacted with a website without like/dislike buttons, the second with a website with only like buttons, and the third with a website with both like and dislike buttons. 
Participants were randomly distributed to one of the three websites in a between-subject design.

We collaborated with a professional panel agency to recruit participants. 
As incentive, participants were awarded digital points, which they could use to get special offers from other companies. 
Participants were above the age of 18 and lived in Germany. 
In a first step, the agency sent their panel members an invitation to participate in the study (_invitation_). 
In this invitation, panel members were asked to participate in a study analyzing the current threat posed by terrorist attacks in Germany.[^1] 
Members who decided to take part were subsequently sent the first questionnaire (_T1_), in which we asked about their sociodemographics, provided more details about the study, and included a registration link for the website. 
Afterward, participants were randomly assigned to one of the three websites. 
After registration participants had the chance to discuss the topic of the terrorism threat in Germany over the course of one week (_field_). 
Subsequently, participants received a follow-up questionnaire in which we collected the self-reported measures (_T2_). 
Measures were collected after and not before the field phase in order not to prime participants or reveal our primary research interest.

@@

We programmed an online website based on the open-source software discourse (https://www.discourse.org/). 
We conducted several pretests with students from the local university to make sure the website had an authentic feel (see Figure \@ref(fig:website)). 
Participants used the website actively: Overall, they spent `r sum(d_raw$time_read, na.rm = TRUE) %>% "/"(60) %>% round(0) %>% format(big.mark = ",")` minutes online, wrote `r sum(d_raw$post_count, na.rm = TRUE) %>% format(big.mark = ",")` comments, and left `r sum(d_raw$reactions, na.rm = TRUE)` popularity cues. 
For an example of communication that took place, see Figure \@ref(fig:comments).

@@

We ran a priori power analyses to determine how many participants to recruit. The power analysis was based on the smallest effect size of interest [SESOI; @lakensEquivalenceTestingPsychological2018]. 
Thus, we defined an effect size that we would consider enough to support our hypotheses. 
Because small effects should be expected when researching aspects of privacy online [e.g., @baruhOnlinePrivacyConcerns2017], with small effects beginning at an effect size of _r_ = .10 [@cohenPowerPrimer1992], we set our SESOI to be r = `r r_sesoi %>% my_round("std")`. 
Our aim was to be able to detect this SESOI with a probability of at least 95%. Using the regular alpha level of 5%, this leads to a minimum sample size of _n_ = `r n_desired %>% format(big.mark = ",")`. 
In the end, we were able to include n = `r n_final` in our analyses (see below). 
This means that our study had a probability (power) of `r power_achieved * 100`% of finding an effect at least as large as _r_ = `r r_sesoi %>% my_round("std")`. 
Put differently, we were able to make reliable inferences about effects at least as big as _r_ = `r r_sensitive %>% my_round("std")`.

We collected a representative sample of the German population in terms of age, sex, and federal state. 
`r n_t1 %>% format(big.mark = ",")` participants completed the survey at T1, `r n_users` participants created a user account on the website,  and `r n_t2` participants completed the survey at T2. 
Using tokens and IP addresses, we connected the data from T1, participants' behavior on the platform, and T2 by means of objective and automated processes.
The data for _n_ = `r n_matched` participants could be matched successfully across all three platforms. 
We excluded _n_ = `r n_speeding` participants who finished the questionnaire at T2 in less then three minutes, which we considered to be unreasonably fast. 
To detect corrupt data, we calculated Cook's distance. 
We excluded `r n_resp_pattern` participants because they provided clear response patterns. 
The final sample included `r n_final` participants.
The sample characteristics at T1 and T2 were as follows: 
T1: Age = `r round(age_t1_m, 0)` years, sex = `r round(male_t1_m, 2)*100`% male, college degree = `r round(college_t1_m, 2)*100`%.
T2: Age = `r round(age_final_m, 0)` years, sex = `r round(male_final_m, 2)*100`% male, college degree = `r round(college_final_m, 2)*100`%.
(One participant did not report his or her sex.) 
<!-- Despite the dropout and although somewhat higher-educated, also T2 can hence be considered a largely representative sample of the German population. -->

>>>>

---
title             : "How Do Like and Dislike Buttons Affect Communication? A Privacy Calculus Approach to Understanding Self-Disclosure Online in a One-Week Field Experiment"
shorttitle        : "Popularity cues and the privacy calculus"
author: 
  - name          : "Author"
    affiliation   : "1"
    corresponding : yes
    address       : "Postal address"
    email         : "my@email.com"
affiliation:
  - id            : "1"
    institution   : "Wilhelm-Wundt-University"
authornote: |
abstract: |
    According to the privacy calculus, both privacy concerns and expected gratifications explain self-disclosure online. So far, however, most findings were based on self-reports, and little is known about whether the privacy calculus can be used to explain observations of actual authentic behavior. Likewise, we still know comparatively little as to whether the privacy calculus can be influenced by the design of online websites, including for example popularity cues such as like and dislike buttons. To answer these questions, we ran a preregistered one-week field experiment. Participants were randomly distributed to three different websites, on which they discussed a current political topic. The final sample consisted of `r knitr::load_cache(label = "load-data", object = "n_matched")` participants. Although the originally preregistered model could not be confirmed, the results showed that the privacy calculus variables predicted a considerable share of actual self-disclosure. The impact of popularity cues was negligible. In conclusion, the results indicate that self-disclosure online can be explained by privacy concerns and specific gratifications. This finding has several implications. For example, it provides further evidence against the privacy paradox.
keywords          : "privacy calculus, self-disclosure, popularity cues, field experiment, structural equation modeling, preregistration"
bibliography      : 
  - [../bibliography/r-references.bib]
  - [../bibliography/references.bib]
floatsintext      : yes
figsintext        : yes
figurelist        : no
tablelist         : no
footnotelist      : no
linenumbers       : yes
mask              : yes
wordcount         : "`r wordcountaddin::word_count('../manuscript/manuscript.Rmd')`"
draft             : no
documentclass     : "apa6"
classoption       : "man"
output            : papaja::apa6_pdf
# output            : papaja::apa6_word
header-includes:
  - \setlength{\parskip}{0em}
  - \raggedbottom
---

@@

Understanding why people disclose personal information online remains a critical question for both society and academic research. 
Originally, self-disclosure online was thought to be mostly erratic.
For example, it was assumed that self-disclosure cannot be predicted by assessing people's personal beliefs, concerns, or standpoints. 
Most prominently, the privacy paradox stated that people self-disclose vast amounts of personal information online *despite* having substantial concerns about their privacy [@barnesPrivacyParadoxSocial2006; @taddickenUsesPrivacyOnline2011].

Somewhat surprisingly, and despite its popularity in the media [@newyorkpublicradioPrivacyParadox2018], the privacy paradox has garnered little empirical support. 
A recent meta-analysis revealed that the correlation between privacy concerns and self-disclosure on SNS is *r* = -.13 [@baruhOnlinePrivacyConcerns2017], which shows that privacy concerns are indeed related to self-disclosure online.

Hence, rather than further pursuing the privacy paradox, a large share of current day research builds on the so-called _privacy-calculus_ [@lauferPrivacyConceptSocial1977], which states that self-disclosure online can be explained---at least partly---by means of expected risks *and* expected benefits [@krasnovaOnlineSocialNetworks2010].
Specifically, by operationalizing expected risks as privacy concerns, several studies have shown that experiencing greater privacy concerns is related to disclosing less information online [@heirmanPredictingAdolescentsDisclosure2013; @koohikamaliInvestigationDynamicModel2019].

@@

Second, current research on the privacy calculus is often criticized for not explicitly focusing on the deliberation process of self-disclosure. 
According to critics [e.g., @knijnenburgDeathPrivacyCalculus2017], showing that concerns and gratifications both correlate with self-disclosure is not evidence for an explicit weighing process of pros and cons. 
We agree.
In this study, we hence explicitly focus on the privacy deliberation process itself. 
Moreover, and on a more general level, we explore the usefulness of further extending the privacy calculus model by adding new variables such as privacy deliberation, trust, and self-efficacy.

Finally, because the privacy calculus does not take place in a vacuum, and because it is often argued that self-disclosure can be easily triggered by external circumstances, we analyze whether the privacy calculus can be affected by the design of a website. 
Specifically, we investigate whether *popularity cues* such as like and dislike buttons have the power to affect the privacy calculus and to foster self-disclosure.

To test our research questions, drawing from a representative sample of the German population we conducted a preregistered online field experiment.
Participants were randomly distributed to one of three different websites, which either included a like button, both a like and a dislike button, or no buttons at all. 
Over the course of one week participants had the chance to discuss a topical issue (i.e., prevention of terrorist attacks in Germany). 
Afterward, they answered a follow-up questionnaire with items measuring the privacy calculus variables.

@@

Being a primary means of regulating privacy [e.g., @masurSituationalPrivacySelfdisclosure2018], self-disclosure is our key variable of interest. 
There are two different understandings of self-disclosure in the literature: The first defines self-disclosure as _deliberate_ acts of sharing truthful information about the self with others [@jourardTransparentSelf1964]. 
The second considers _all_ acts of sharing information---be they active or passive, deliberate or unwitting---as self-disclosure, because each piece of information shared allows for meaningful inferences about a person [@watzlawickPragmaticsHumanCommunication2011]. 
In this paper we follow the latter approach, not least because the recent years have vividly illustrated how easy it is to derive personal insights simply by analyzing exchanged communication [@kosinskiPrivateTraitsAttributes2013]. 
Moreover, independent from which position one chooses to adopt, it is possible to differentiate the content of self-disclosure into three different dimensions: breadth (i.e., number of topics covered), depth (i.e., intimacy of topics covered), and length (i.e., quantity of disclosure) [@omarzuDisclosureDecisionModel2000]. 
In this study we mainly focus on communication quantity, as we consider communication quantity to be a necessary precondition and hence valid proxy for self-disclosure.

Privacy concerns have been defined as follows: 
"Concerns about online privacy represent how much an individual is motivated to focus on his or her control over a voluntary withdrawal from other people or societal institutions on the Internet, accompanied by an uneasy feeling that his or her privacy might be threatened" [AUTHOR]. 
Previous research has found that people who are more concerned about their privacy than others are less likely to share personal information [@baruhOnlinePrivacyConcerns2017; @dienlinPrivacyParadoxRelic2015; @heirmanPredictingAdolescentsDisclosure2013; @koohikamaliInvestigationDynamicModel2019].

@@

Although privacy concerns are related to self-disclosure, one can make the case that since most studies in the literature report only small effects there should also be additional meaningful factors that contribute to explaining self-disclosure. 
Most prominently, it has been argued that people trade a loss of privacy for a gain in gratifications such as social capital, entertainment, information, or self-presentation [@ellisonNegotiatingPrivacyConcerns2011; @taddickenUsesPrivacyOnline2011]. 
By now, a large body of research has found support for this hypothesis [e.g., @krasnovaOnlineSocialNetworks2010; @minHowArePeople2015; @trepteCrossculturalPerspectivePrivacy2017].

@@

As mentioned above, there is still a shortage of studies that explicitly analyze the decision process behind the disclosing of information, although this point of criticism has been leveled several times [@knijnenburgDeathPrivacyCalculus2017] and although other fields such as behavioral economics have long focused on the underlying problem [@zhuPrivacyCalculusIts2017]. 
This criticism is justified.
The observation that privacy concerns and expected gratifications are related to self-disclosure is by itself not a sufficient condition for an explicit weighing process.
Hence, research on the privacy calculus would benefit significantly from analyzing this decision process explicitly. 
Building on @omarzuDisclosureDecisionModel2000 and @altmanPrivacyConceptualAnalysis1976, we hence address a novel concept that might best be termed _privacy deliberation_, which captures the extent to which individual people explicitly compare positive and negative potential outcomes before communicating with others.

On the one hand, it seems plausible that deliberating about one's privacy would dampen subsequent self-disclosure, because refraining from regular communication---the primary means of connecting with others---requires at least a minimum of active and hence deliberate restraint. 
On the other hand, deliberating about one's privacy might also increase self-disclosure, because a person concerned about his or her privacy might arrive at the conclusion that in this situation self-disclosure is not only appropriate but expedient. 
In light of the lack of empirical studies and the plausibility of both effects, we formulate the following research question:

@@

Several attempts have already been made to expand the privacy calculus [@dinevExtendedPrivacyCalculus2006], introducing additional variables such as self-efficacy or trust. 
Self-efficacy in the context of the privacy calculus captures whether people believe in their own capability to implement particular privacy behaviors in the future [@dienlinExtendedPrivacyCalculus2016].
These privacy behaviors refer to either self-_disclosure_ (e.g., publishing a blog post) or self-_withdrawal_ (e.g., deleting inappropriate content).
People who report more privacy self-efficacy also engage in more self-withdrawal  [@chenRevisitingPrivacyParadox2018].
In light of our focus on active communication, in this study we investigate the influence of self-_disclosure_ self-efficacy.

@@

Trust can be conceptualized in two different ways [@gefenTrustTAMOnline2003]: 
It either captures "_specific_ beliefs dealing primarily with the integrity, benevolence, and ability of another party" [@gefenTrustTAMOnline2003, p. 55, emphasis added] or a "_general_ belief that another party can be trusted" [@gefenTrustTAMOnline2003, p.55, emphasis added]. 
Whereas specific trust focuses on the causes of trust, general trust emphasized the experience of trust.
@gefenTrustTAMOnline2003 prioritize specific trust (p. 60). 
In the online context, it is also important to differentiate among several _targets_ of trust [@sollnerWhyDifferentTrust2016]. 
Potential targets include (a) the information system, (b) the provider, (c) the Internet, and (d) the community of other users [@sollnerWhyDifferentTrust2016]. 
Trust plays a key role in online communication [@metzgerPrivacyTrustDisclosure2004]. 
For example, people who put more trust in the providers of networks also disclose more personal information [@liEmpiricalStudiesOnline2011].

@@

What is the effect of the communication context on the privacy calculus and on self-disclosure? 
First, it has often been noted that researchers should not exclusively focus on specific features of particular websites, for features are prone to change and quickly become obsolete [@foxDistinguishingTechnologiesSocial2017]. 
Instead, it has been suggested to prioritize underlying latent structures, for example by analyzing so-called affordances [@ellisonSocialNetworkSite2015; @foxDistinguishingTechnologiesSocial2017]. 
The concept of affordances was developed by @gibsonEcologicalApproachVisual2015, who argued that it is not the _objective features_ of objects that determine behavior.
Instead, more important are the _subjective perceptions_. 
Affordances are a mental representation of how a given entity might be used; as such, they are by definition subjective. 
There is an ongoing debate on what exactly defines an affordance [@evansExplicatingAffordancesConceptual2017]. 
For example, whereas @evansExplicatingAffordancesConceptual2017 propose three affordances for mediated communication (i.e., anonymity, persistence, and visibility), @foxDistinguishingTechnologiesSocial2017 suggest 10 affordances for SNSs alone (i.e., accessibility, bandwidth, social presence, privacy, network association, personalization, persistence, editability, conversation control, and anonymity).

As the privacy calculus states that both benefits and costs determine behavior, we suggest that popularity cues such as like and dislike buttons, which are categorized as "paralinguistic digital affordances" [@carrPredictingThresholdPerceived2018, p. 142], perfectly capture potential benefits and costs. 
The like button is positive. 
It expresses an endorsement, a compliment, a reward [@carrPredictingThresholdPerceived2018; @sumnerFunctionalApproachFacebook2017]. 
However, communication online is also often characterized by negative and critical debates [@ziegeleDynamicsOnlineNews2017]. 
As the dislike button is a major means of downgrading content it is negative and represents the risk factor of the privacy calculus well. 
In fact, its stark negative effect might also explain why to date only a handful of major websites have implemented it (e.g., youtube, reddit, or stackexchange).

Paralinguistic digital affordances and specifically popularity cues have been shown to impact behavior [@kramerMasteringChallengeBalancing2020; @treptePrivacyCalculusContextualized2020].
For example, a large-scale field experiment in which 101,281 comments were analyzed found that comments with dislikes were more likely to receive further dislikes [@muchnikSocialInfluenceBias2013]. 
@stroudRecommendRespectAltering2017 demonstrated that when users disagreed with a post, they were more likely to click on a button labeled _respect_ compared to a button labeled _like_.

In this vein, it seems plausible that popularity cues might also impact the privacy calculus [@kramerMasteringChallengeBalancing2020].
First, popularity cues could serve as a means of reward and punishment, affecting behavior via instrumental conditioning [@skinnerScienceHumanBehavior2014]. 
Being complimented with a like should encourage future self-disclosure, while being punished with a dislike should inhibit disclosure. 
Similarly, like buttons imply being able to garner positive feedback, so implementing a like-button---similar to a compliment in the offline world---might leverage gratifications. 
Implementing popularity cues might also bring people to more actively deliberate about whether or not it is actually worthwhile to disclose information. 
If both like and dislike buttons are present, privacy deliberation should increase even further. 
Finally, because people who are more concerned about their privacy are also more shy and risk averse [@dienlinPsychologyPrivacyAnalyzing2017], implementation of the dislike button should both stir privacy concerns and stifle self-disclosure.

@@

For a simplified overview of our theoretical model, see Figure \@ref(fig:model).

@@

The online supplementary material (OSM) of this study includes the data, research materials, analyses scripts, and a reproducible version of this manuscript (see https://osf.io/hcqat/?view_only=5db35868738d40609b11e58cc343a9b0).
We preregistered the study using the registration form _OSF Prereg_, which includes the hypotheses, sample size, research materials, analyses, and exclusion criteria (see https://osf.io/a6tzc/?view_only=5d0ef9fe5e1745878cd1b19273cdf859). 
We needed to change our pre-defined plan in some cases. 
For a full account of all changes, see OSM. 
New analyses that were not preregistered appear in the section on exploratory analyses.

@@

The study was designed as an online field experiment with three different groups. 
The first group used a website without like/dislike buttons, the second a website with only like buttons, and the third a website with both like and dislike buttons. 
Participants were randomly distributed to one of the three websites in a between-subject design.

We collaborated with a professional panel agency to recruit participants. 
As incentive, participants were awarded digital points, which they could use to get special offers from other companies. 
Participants were above the age of 18 and lived in Germany. 
In a first step, the agency sent its panel members an invitation to participate in the study (_invitation_). 
In this invitation, panel members were asked to participate in a study analyzing the current threat posed by terrorist attacks in Germany.[^1] 
Members who decided to take part were subsequently sent the first questionnaire (_T1_), in which we asked about their sociodemographics, provided more details about the study, and included a registration link for the website. 
Afterward, participants were randomly assigned to one of the three websites. 
After registration was completed, participants could discuss the topic of the terrorism threat in Germany over the course of one week (_field_). 
Subsequently, participants received a follow-up questionnaire in which the self-reported measures were collected (_T2_). 
Measures were collected after and not before the field phase in order not to prime participants or reveal our primary research interest.

@@

We programmed an online website based on the open-source software _discourse_ (https://www.discourse.org/). 
We conducted several pretests with students from the local university to make sure the website had an authentic feel (see Figure \@ref(fig:website)). 
Participants used the website actively: Overall, they spent `r sum(d_raw$time_read, na.rm = TRUE) %>% "/"(60) %>% round(0) %>% format(big.mark = ",")` minutes online, wrote `r sum(d_raw$post_count, na.rm = TRUE) %>% format(big.mark = ",")` comments, and left `r sum(d_raw$reactions, na.rm = TRUE)` popularity cues. 
For an example of communication that took place, see Figure \@ref(fig:comments).

@@

We ran a priori power analyses to determine how many participants to recruit. The power analysis was based on a smallest effect size of interest [SESOI; @lakensEquivalenceTestingPsychological2018]. 
In other words, we defined a minimum effect size that we would consider sufficiently large enough to support our hypotheses. 
Because small effects should be expected when researching aspects of privacy online [e.g., @baruhOnlinePrivacyConcerns2017], with small effects beginning at an effect size of _r_ = .10 [@cohenPowerPrimer1992], we set our SESOI to be _r_ = `r r_sesoi %>% my_round("std")`. 
Our aim was to be able to detect this SESOI with a probability of at least 95%. Using the regular alpha level of 5%, this leads to a minimum sample size of _n_ = `r n_desired %>% format(big.mark = ",")`. 
In the end, we were able to include n = `r n_final` in our analyses (see below). 
This means that our study had a probability (power) of `r power_achieved * 100`% to find an effect at least as large as _r_ = `r r_sesoi %>% my_round("std")`. 
Put differently, we were able to make reliable inferences about effects at least as big as _r_ = `r r_sensitive %>% my_round("std")`.

We collected a representative sample of the German population in terms of age, sex, and federal state. 
`r n_t1 %>% format(big.mark = ",")` participants completed the survey at T1, `r n_users` participants created a user account on the website,  and `r n_t2` participants completed the survey at T2. 
Using tokens and IP addresses, we connected the data from T1, participants' behavior on the platform, and T2 by means of objective and automated processes.
The data for _n_ = `r n_matched` participants could be matched successfully across all three platforms. 
We excluded _n_ = `r n_speeding` participants who finished the questionnaire at T2 in less then three minutes, which we considered to be unreasonably fast. 
To detect corrupt data, we calculated Cook's distance. 
We excluded `r n_resp_pattern` participants because they provided clear response patterns. 
The final sample included `r n_final` participants.
The sample characteristics at T1 and T2 were as follows: 
T1: Age = `r round(age_t1_m, 0)` years, sex = `r round(male_t1_m, 2)*100`% male, college degree = `r round(college_t1_m, 2)*100`%.
T2: Age = `r round(age_final_m, 0)` years, sex = `r round(male_final_m, 2)*100`% male, college degree = `r round(college_final_m * 100, 0)`%.
One participant did not report his or her sex.
<!-- Despite the dropout and although somewhat higher-educated, also T2 can hence be considered a largely representative sample of the German population. -->
